---
title: "Joining Assignment"
author: "Your Name Here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

```

# Load the datasets

```{r}

counties_demographics <- readRDS(here("data", "counties_demographics.rds"))
counties_elexresults <- readRDS(here("data", "counties_elexresults.rds"))
counties_rural_urban_suburban <- readRDS(here("data", "counties_rural_urban_suburban.rds"))



```

# Exercises

1)  

Join together the election results and urban/suburban/rural tables above, and then use the merged table to:

-   filter for just Virginia

-   group_by and summarize to calculate the mean pct of Trump 2020 vote (`pct_r_2020`) by urban/rural categories (`rural_urban`)

```{r}

elex_rural_urb_suburb <- left_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid")

elex_rural_urb_suburb %>% 
  filter(state_code == "VA") %>% 
  group_by(rural_urban) %>%
  summarise(mean_trump_pct = mean(pct_r_2020, na.rm = TRUE))

```

2)  

Using the joined election results / urban/suburban/rural table created above:

-   filter for just Virginia

-   count up how many counties were won by party in 2020 (`winner_2020`) by urban/rural categories

```{r}

elex_rural_urb_suburb %>% 
  filter(state_code == "VA") %>% 
  group_by(winner_2020) %>% 
  summarise(counties_won_count = n())


```

3)  

Join together the election results and demographics tables above, and then use the joined table to:

-   filter for just Florida

-   group/summarize to calculate the mean pct of Trump 2020 vote in the counties above or below national average for college graduates (`pct_ed_college_all_abovebelow_natl`)

```{r}

elex_demographics <- left_join(counties_elexresults, counties_demographics, by = "geoid")

elex_demographics %>% 
  filter(state_code == "FL") %>% 
  group_by(pct_ed_college_all_abovebelow_natl) %>%
  summarise(mean_trump_pct = mean(pct_r_2020.x, na.rm = TRUE))


```

4)  

Using the joined table you made above:

-   group/summarize to calculate the mean pct of Trump 2020 vote by counties above or below the national average for college graduates (`pct_ed_college_all_abovebelow_natl`).

    Note we're looking for the entire country here.

```{r}

elex_demographics %>%
  group_by(pct_ed_college_all_abovebelow_natl) %>%
  summarise(mean_trump_pct = mean(pct_r_2020.x, na.rm = TRUE))


```

Now using the joined datasets above, come up with two groupings you find interesting or newsworthy in a particular state, and replicate the steps above them them instead.

5)  
comparing the percentage of military veterans that voted for Trump in the state of MA (which is known
to be a very blue state) to the rest of the country

```{r}

elex_demographics %>%
  filter(state_code == "MA") %>% 
  group_by(pct_mil_veteran_abovebelow_natl) %>%
  summarise(mean_trump_pct = mean(pct_r_2020.x, na.rm = TRUE))

elex_demographics %>%
  group_by(pct_mil_veteran_abovebelow_natl) %>%
  summarise(mean_trump_pct = mean(pct_r_2020.x, na.rm = TRUE))

```

6)  
comparing the mean percentage of trump votes to the total votes in the state of NY in the 2020 election

```{r}

elex_rural_urb_suburb %>% 
  filter(state_code == "NY") %>% 
  group_by(rural_urban, totalvotes_2020) %>%
  summarise(mean_trump_pct = mean(pct_r_2020, na.rm = TRUE))

```

7)  Extra credit

Up until now we've joined two tables together. But we actually have three tables in this project. How might we join all three together into a single table, so all the measures for each county are combined in one place? (Hint: you can do this is multiple steps, doesn't have to be one step only.)

```{r}

elex_dems_2 <- full_join(counties_elexresults, counties_demographics, by = "geoid")

elex_dems_full <- full_join(elex_dems_2, counties_rural_urban_suburban, by = "geoid")


```
